### 2020-08-27 supplement/supplement.tex

\documentclass[aip, numerical, preprint]{revtex4-2}

\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{siunitx}

\usepackage[colorlinks=true,urlcolor=blue,citecolor=blue]{hyperref}
\usepackage{xcolor}
\usepackage{enumitem}
\setlist{nosep}

\begin{document}
\title{Supplementary Materials: Introductory Physics Labs: A Tale of Two Transformations}

% \author{Steven F.\ Wolf} \affiliation{East Carolina University Department of Physics 1000 E.\
% 5\textsuperscript{th} street, Greenville, NC 27858 USA} \author{Mark W.\ Sprague}
% \affiliation{East Carolina University Department of Physics 1000 E.\ 5\textsuperscript{th}
% street, Greenville, NC 27858 USA}

\author{First Author} \affiliation{Masked Institution}

\author{Second Author} \affiliation{Masked Institution}

\date{\today}

\maketitle

\section{Curriculum development timeline}
Talk about the curriculum development, and how departmental culture around the labs changed as
a part of this transformation.

\section{Face-to-face implementation}

Activities and basic equipment.

Assessment practices and discussion of results.

\section{Online implementation}

Curricular adaptation.

Kit contents.

\section{Introduction}
There are many challenges that physics faculty face when teaching introductory labs. One of
these challenges is engaging students in authentic science practices.  Another has been
highlighted given the current global pandemic---specifically how to engage students in our
laboratory courses while maintaining appropriate social distancing and hygiene standards. We
have chosen to answer these challenges by transforming our labs\ldots\ twice. We discuss the
rationale behind the first transformation to a practice-focused curriculum.  However, when
March 2020 rolled around we needed to transform our labs again, this time to accommodate online
learning. This paper discusses three chief questions: ``Why did we change?'' ``What are we
doing to engage students in science practices?'' and ``How did we make all of this work
online?''

Scientific truth is discovered through the practice of doing science.  \citeauthor{DBERreport}
state, ``Without [scientific] investigative practices, there would be no new scientific and
engineering knowledge.''\citep{DBERreport} Too often, our laboratory courses do not
meaningfully engage students in scientific practices, instead reproducing classical
experiments.\citep{PCAST12} This is part of a long-standing tradition of calling for authentic
engagement in scientific practices which pre-dates Lorentz transformations.\citep{AAAS1881}
Indeed, back in \citeyear{AAAS1881}, the AAAS stated our goals quite succinctly, ``[Typical
courses in t]he sciences\ldots are not made the means of cultivating the observing powers,
stimulating inquiry, exercising the judgment in weighing evidence, nor of forming independent
habits of thought.''\citep{AAAS1881}

The physics labs at our institution in the Fall of 2016 very much mirrored the national
narrative.  The lab curriculum had been largely unchanged for about 20 years, and followed the
traditional cookbook lab paradigm.  Moreover, the equipment commonly used by students was
outdated---triple beam balances and hand-drawn graphs were standard.  In the same year, as part
of an interdisciplinary team interested in engaging students in scientific practices as well as
assessing those scientific practices, we secured funding, both internally and from the NSF, to
change this status quo in physics, biology, and chemistry.  All three departments transformed
their labs using a pedagogical framework entitled, Argument Driven Inquiry
(ADI),\citep{Sampson2011,Walker2011,Walker2016} which has been specifically designed to engage
students in authentic science practices. We will focus on the physics transformation in this
paper.

We were satisfied with our progress when March 2020 rolled around, and with it, the mandatory
shut-down due to the global COVID-19 pandemic.  With students off campus, we had to abruptly
shift our delivery to an online format and quickly realized that some of the activities that we
utilized in the face-to-face sections had to be adapted to online use.  Going forward, we have
further adapted what we are doing so that students can still participate in a hands-on
laboratory experience in an online laboratory course.  We hope these changes will be meaningful
long-term as a lack of online labs is also a barrier for distance education students'
completion of a degree.  So online lab curricula would fill an institutional need.  In the next
sections, we will discuss our institutional and instructional contexts and describe each
transformation in more detail.

\section{Instructional Context}

Regional State University is a public doctoral granting university in a rural, economically
depressed part of our state.  The university mission is focused on regional transformation and
service to this region.  The total combined student population in Fall 2019 was 28,651 largely
comprised of undergraduates (23,081).Our physics labs (Physics 1 lab and Physics 2 lab) are
each 1 credit courses that meet for 2 hours per week. These courses serve both of our
Calculus-based and Algebra-based physics lecture courses.  Most $(\sim 75\%)$ of the students
who take our lab course are in (or have taken) the Algebra-based physics lecture course. It
also fulfills the general education requirement for science lab courses at our university.
Table \ref{tab:gender} compares the gender of students in these lab courses compared to the
university population.  We note that the gender distribution of students in the Physics 2 lab
generally matches the university population, while the Physics 1 lab skews more heavily male
than the university population.  We also compare the race of students in these contexts in
Table \ref{tab:race}.  The racial and ethnic profile of students in our lab courses is similar
to the university.

Students who take these labs are most often science majors, especially in the health sciences
(e.g., Biology, Exercise Science), but this lab is also taken by our physics majors.  In Fall
2019 we had 409 students in the Physics 1 lab, and in Spring 2020 (at Census Day, before the
pandemic changed enrollments) we had 256 students in the Physics 2 lab.  In both labs, our
enrollment is capped at 22 students, due to the constraints of our laboratory classroom.

Labs are supervised by Author 2, but each section is run by Graduate GTAs (GTAs). The
transformed curriculum was jointly written by both of the authors.  GTA training was greatly
enhanced when the new lab curriculum was put in place.  Both authors, along with colleagues in
Biology and Chemistry, run a training for all GTAs in Biology, Chemistry, Geology, and Physics
as these disciplines are all using the same curricular format for their labs.  Author 2 runs a
weekly prep meeting with the GTAs, and SFW and the research team have supervised various
aspects of the transformation especially important for research such as curricular
implementation \citep{SmithJoyner2020} and assessment grading practices \citep{Wolf2019}.

\begin{table}
  \centering
  \begin{tabular}{lrrr}
    \hline \hline
                &\multicolumn{2}{c}{Course}\\
    Gender	&Physics 1 &Physics 2 &University\\
    \hline
    Female	&41.8\%	   &60.9\%    &57.1\%\\
    Male	&58.2\%	   &39.1\%    &42.9\%\\
    \hline \hline
  \end{tabular}
  \caption{Gender breakdown of students in the Fall 2019 of the Physics 1 lab, the Spring 2020
    Physics 1 lab, and the undergraduate poplulation of the university at large in Fall 2019.
    Note: The university only collected binary gender data.}
  \label{tab:gender}
\end{table}

\begin{table}
  \centering
  \begin{tabular}{lrrr}
    \hline\hline
                                            &\multicolumn{2}{c}{Course}\\
    Race/Ethnicity                          &Physics 1  &Physics 2 &University\\
    \hline
    White                                   &60.8\%     &60.8\%    &65.3\%\\
    Black or African American               &17.1\%     &15.5\%    &16.3\%\\
    Hispanic                                &7.5\%      &8.7\%     &7.5\%\\
    Two or More Races                       &7.3\%      &5.3\%     &3.8\%\\
    Unknown                                 &2.9\%      &3.0\%     &3.4\%\\
    Asian                                   &4.0\%      &6.4\%     &2.4\%\\
    American Indian or Alaska Native        &.	        &.         &0.6\%\\
    Non-Resident Alien	                    &0.4\%      &0.4\%     &0.6\%\\
    Native Hawaiian/Other Pacific Islander  &.	        &.         &0.1\%\\
    \hline\hline
  \end{tabular}
  \caption{Race/Ethnicity breakdown of students inthe Fall 2019 semester of the Physics 1 lab,
    the Spring 2020 semester of the Physics 2 lab, and the undergraduate population of the
    university at large in Fall 2019.}
  \label{tab:race}
\end{table}

\section{Transformation \#1: Argument Driven Inquiry}
Argument Driven Inquiry (ADI) is a pedagogical method that was developed to make science labs
in school better match what actually transpires in authentic research settings
\citep{Walker2011,Sampson2011,Walker2016}.  In traditional labs, we often ask students to
answer questions, or measure quantities, that are well-known using procedures that have been
refined over time.  For example, in physics we often have students measure $g$ using a
well-known method (such as a spark timer).  In ADI labs, we give students a question to answer
(e.g., ``Does the force a fan exerts on a cart depend on the mass of the cart?''), allow them
to design a procedure for an experiment that will answer that question, collect and analyze
their data, and make claims based on their data.  We accomplish this by having students engage
in a three-part cycle:
\begin{itemize}
  \item Pre-Lab - Introducing the context
  \item Proposal Development and Data Collection
  \item Argumentation and Peer Review
\end{itemize}
In the sections below, we will comment on each part of the cycle, focusing on both students and
GTAs.

\subsection{Session 1: Pre-Lab}
Part of the goal of any laboratory is to help students develop a familiarity with certain tools
and equipment common to the lab.  The purpose of the pre-lab reading and activity is to give
students familiarity with the equipment they will be using.  For the investigation described
previously, ``Does the force a fan exerts on a cart depend on the mass of the cart?''  we need
students to become familiar with a few things, such as what we mean by a cart, and give them
some tools that they need to be able to answer this question.  As a part of the pre-lab
reading, students are given information about linear regression---both how to run it in Excel,
and how to interpret the output, and how we can estimate instantaneous velocity given position
vs.\ time data. For this experiment, we give them a cart (leaving the fan off for now) and have
them measure the acceleration of that cart down a ramp using an ultrasonic motion detector to
collect position vs.\ time data for this motion.  Students work in pairs to do this work, and
turn in a brief summary of this work on the course management system.  For this investigation,
it is the first time that they create graphs, so we set norms for graphing such as helping them
decide which variable to put on the x-axis and the y-axis as well as including units.  Pre-lab
assignments are designed to be graded in less than 1 minute/paper based on a rubric we provide
to both students and GTAs.

\subsection{Session 2: Proposal Development and Data Collection}
During this session, students work in groups of 4 to develop a proposal, get it approved by a
GTA, and then collect and analyze their data.  We have a format for proposal development that
asks students to link the scientific concept being studied---in this example case, Newton's
Second Law---to the data that they will collect.  Students also need to propose a plan for
analyzing their data and minimizing potential sources of error (such as friction).  Once
students have an approved proposal, they begin collecting and analyzing their data.  This may
lead them to refine or revise their method (which they are encouraged to do).  Grading for the
GTAs this week is minimal (full credit for completing the proposal).  However, this is the most
critical week's for GTA training.  GTAs have to be clear that the goal is to allow a diverse
number of ways of collecting/analyzing this data.  How much does the mass of the cart need to
vary?  How many different cart masses do the students need to take?  How should students deal
with friction? All of these are questions that we want the \textit{students} to answer.  And in
some sense, we will be disappointed if all of the students come up with the same answer.
Experimental methods certainly have an impact on experimental results.

\subsection{Session 3: Argumentation and Peer Review}
During this session, the students begin by sharing their results.  Students finish analyzing
their data during the prior week, and prepare a ``poster'' on a whiteboard.  One person from
the group stays at the whiteboard while the other group members (travelers) go to other groups.
The travelers learn what other groups did, and compare to their results in a structured
argumentation setting.  Groups spend about 3--5 minutes at a poster, and rotate so that they
see at least 3 other posters.  After the travelers return, critically, the groups spend time
discussing what they saw and reflecting on their own results.  What is exciting is that
sometimes, these interactions actually drive groups to change their claims \citep{Walker2019}.

Each student writes a shortened lab report consisting of three sections: an introduction to the
scientific concept, a description of the experimental method or procedure, and a discussion
section laying out their results and the evidence and justification supporting their results.
The authors have created different peer review calibration videos for the students to watch so
they can look for the same things that we are going to be grading them on when we get their
final reports.  Students complete their lab report and submit it by the end of the day.  Then,
students take part in double-confidential peer review.  We facilitate this online.  They read
and reply to two different lab reports.  Finally, they get their feedback and use this to
revise their own lab report.  At the end of this week, students turn in their revised,
peer-reviewed final lab report.  This is the heaviest week of GTA grading.  We provide a rubric
for lab report grading, and each lab report generally takes about 3-5 minutes to grade.

\section{Transformation \#2: Online Adaptation}

\subsection{Sudden Transition to Online Instruction--Spring 2020}

In March 2020 the COVID-19 pandemic forced most universities, including ours, to move all their
classes online. Our General Physics I and II laboratories had completed two out of four full
investigations and the pre-lab for the third investigation face-to-face. We were forced to find
a way to engage students in an online format while preserving the nature of the ADI laboratory
experience. The face-to-face activities that we moved online were the Investigation 3 Proposal
and Argumentation and the Investigation 4 Pre-Lab, Proposal, and Argumentation. In addition, we
gave our laboratory practical exam online. We required student investigation groups to find a
method for online collaboration in which everyone in the group could participate. Some groups
used Cisco WebEx (video interaction platform licensed by our university) sessions, some used a
group chat, and others used FaceTime or other online communication applications.

During the first week of online classes each group produced their proposal and posted a
proposal form on a Canvas Discussion for approval. The GTA reviewed proposals and provided
feedback or approval. Groups used the GTA feedback to revise their proposals until they were
approved. Most proposals obtained approval after two or three revisions, but some required as
many as seven revisions to obtain approval. The GTAs and students found communication about
proposals much more difficult online than in face-to-face classes. When the GTA approved a
group's proposal, they assigned the group a data set for the investigation, and the group began
its measurements and analysis. We provided measurements to the students in the most raw form
possible to require them to make decisions about data collection and analysis.

Investigation 3 in the General Physics Laboratory I course was a study of the periodic motion
of a mass hanging from a spring in which the students were asked to determine when the spring's
mass must be considered as a contribution to the period. Before the transition to online
classes the students had completed a pre-lab activity in which they measured the period of a
mass on a spring. For this investigation we provided videos of various masses oscillating on
springs (100 oscillations each for 10 different masses and for the spring oscillating with no
mass) and photographs of each spring and each mass on a balance. We posted video and photograph
sets for six different springs so groups in the same lab section would each have different
springs to study.

Investigation 3 for the General Physics Laboratory II course was a study of light diffraction
in which the students were asked to determine whether hairs from two individuals had the same
diameter. Before the transition to online classes, the students had completed a pre-lab
activity in which they determined the width of a single slit by measuring the diffraction
pattern. We collected hair samples from several people and posted photographs of the
diffraction pattens of the hairs. Each photograph had a ruler at the bottom for the students to
use as a length scale. We also provided photographs of the positions of the holder and screen
on the optics bench. We gave the students a tutorial on using the \emph{ImageJ}
application\citep{schrasetal12} to measure distances in a photograph.

Each group was required to complete its analysis and create a three-slide presentation for the
argumentation session the following week. The first slide was a description of their
measurements. Te second slide was a presentation of the results, including a graph or table,
and the third slide was their argument, based on their result.

The argumentation session was held in a Cisco WebEx session during the lab session the week
after the proposal session. One member of each group gave the presentation, which was followed
by questions. Students received credit for giving presentations, asking meaningful questions,
and responding to questions.  Following the argumentation session students submitted individual
draft reports, peer-reviewed each other's drafts, and submitted final reports in the same
manner used for the face-to-face investigations.

Investigation 4 in the General Physics Laboratory I course was about collisions, and students
were asked to determine whether a collision between two marbles was elastic. Our original plans
were to have the students video marble collisions in lab and analyze them using the
\emph{Tracker} application\citep{bro2009}, which is installed on the computers in the
laboratory. When the labs went online, we provided the students with several videos of marble
collisions and asked the students to install \emph{Tracker} on their computers for analysis.
We adapted the Investigation 4 Pre-Lab assignment so students could perform them on their own
computers. The originally-planned pre-lab was an analysis of a video using \emph{Tracker}. This
activity required only a few changes from the face-to-face pre-lab assignment. We conducted the
proposal session as described for Investigation 3 and assigned each group one of six videos of
colliding marbles to analyze along with mass measurements of the marbles in the videos. The
argumentation session and the remainder of the investigation were conducted in the same manner
as Investigation 3.

Investigation 4 in the General Physics Laboratory II course was a study of radioactive
decay. The pre-lab for the face-to-face course is a simulation of radioactive decay using dice
in which the students roll several dice and remove all the dice with one dot showing. We wrote
a GlowScript\citep{glowscript} program to ``roll'' randomized virtual dice so they could
perform the same activity using this simulation on their computers. For the investigation the
students were asked to determine which isotope was most common in the nuclear decay of a copper
disk that had been exposed to low-energy neutron radiation. We measured radiation counts for
several disks and also background levels and provided students with \SI{30}{s} counts vs.\ time
in CSV files for analysis. As in the other course, the investigation was conducted in the same
manner as Investigation 3.

We administered the lab practical exams for both courses in Canvas using GlowScript simulations
embedded in Canvas assignments. Students made measurements on the simulation and used their
results to make an argument answering a guiding question.

We encountered many problems with the move to online laboratories. Our students had not
registered for an online class, and many were not prepared for the sudden transition from
face-to-face to online classes. Many students could not or did not attend the online WebEx
sessions or participate with their assigned groups. Some students had to get jobs when they
returned home, and many students did not have access to high-speed internet. We removed
non-participating students from groups and gave them an opportunity to make up their missed
work asynchronously. Less than \SI{50}{\percent} of the students in the make-up groups
completed their work. We resorted to dropping the lowest investigation for the course.

Some students in the course did not have access to computers capable of running \emph{Tracker}
or \emph{ImageJ}, both of which run on Windows, Macintosh, or Linux computers but not
Chromebooks or mobile devices such as smartphones or tablets. We discovered
\emph{jsTrack}\citep{jstrack}, an online Javascript web application for video analysis that
runs on most computers including Chromebooks. Students using mobile devices were not able to
use \emph{jsTrack} either.

\subsection{Fully Online Laboratories--Fall 2020}

\begin{table}
  \caption{\label{tab: 1251 fall} General Physics Laboratory I investigations for Fall 2020
    block schedule.}
  \begin{ruledtabular}
    \begin{tabular}{ccp{28em}}
      Investigation & Topic & Guiding Question\\
      \hline
      1 & 1-D kinematics & Does a ball rolling on an incline have the same acceleration on the way up as it does on the way down? \\
      2 & Periodic motion & At what nut position is the period of the physical pendulum equal to \SI{1.30}{s}? \\
      3 & 2-D collisions & Is the collision between two marbles elastic?
    \end{tabular}
  \end{ruledtabular}
\end{table}

We decided to hold our introductory physics laboratory courses online in the Fall 2020 semester
in order to preserve the group class interaction aspects of ADI, which would be difficult under
the social distancing requirements in place due to the pandemic.\citep{mclber20} Also our
teaching laboratories would have to operate at half-capacity face-to-face to maintain social
distancing, preventing us from offering the courses to the necessary number of students.  In
addition to the social distancing requirements for face-to-face class meetings, our university
adopted an eight-week block schedule, with the second block ending before the Thanksgiving
holiday. Half of the Fall 2020 courses were scheduled for the first eight-week block, and half
of the courses were scheduled for the second block. Course mapping between the originally
scheduled 14-week semester and the two eight-week blocks was based on the originally scheduled
class meeting time. In the block schedule, the one semester-hour lab courses have two two-hour
meetings per week. We determined that, even though there were enough lab meetings for the
synchronous activities of four full investigations, there was not enough time between lab
meetings for the asynchronous components and timely grading for four investigations. We reduced
the number of investigations to three and added an additional pre-lab activity to each
investigation. The topics for the three investigations and the guiding questions are in Table
\ref{tab: 1251 fall} for General Physics Laboratory I and in Table \ref{tab: 1261 fall} for
General Physics Laboratory II.

\begin{table}
  \caption{\label{tab: 1261 fall} General Physics Laboratory II investigations for Fall 2020
    block schedule.}
  \begin{ruledtabular}
    \begin{tabular}{ccp{26em}}
      Investigation & Topic & Guiding Question\\
      \hline
      1 & Current and resistance & Does a light bulb behave like a resistor? \\
      2 & Time varying circuits & Do two of the lab kit capacitors have the same capacitance? \\
      3 & Diffraction of light & Are hairs from different people the same diameter?
    \end{tabular}
  \end{ruledtabular}
\end{table}

We informed the students before the course began that internet connectivity was required and
that they must have access to a computer capable of running \emph{Tracker} (General Physics
Laboratory I) or \emph{ImageJ} (General Physics Laboratory II). We also included these
statements in the course syllabus. Although the courses were online, most of the students were
on campus allowing them to access campus computer laboratories if they did on a computer that
met the course requirements.

\begin{table}
  \caption{\label{tab: 1251 lab kit} General Physics Laboratory I lab kit contents.}
  \begin{tabular}{clc}
    \hline\hline
    Quantity & Item & Investigation(s)\\
    \hline
    1 & Protractor & 1 \\
    2 & \SI{25}{mm} marble & 1, 3 \\
    1 & Tape measure with cm scale & 1, 2, 3\\
    1 & \SI{0.6}{m} threded rod & 2 \\
    1 & Eye nut & 2 \\
    3 & Nuts & 2 \\
    1 & Door hook & 2 \\
    1 & \SI{1}{m} string & 2 \\
    1 & \SI{16}{mm} marble & 3 \\
    \hline\hline
  \end{tabular}
\end{table}

We developed lab kits with supplies that allowed the students to perform the investigations
outside the teaching laboratory. We purchased the lab kit items in collaboration with our
campus bookstore, and the students purchased the lab kits from the bookstore. We ordered lab
kit items in bulk and where possible directly from manufacturers to reduce the costs of the
items. Each General Physics Laboratory I kit costs \textcolor{red}{$\sim$}\$25.00, and each
General Physics Laboratory II kit costs \$39.00. Table \ref{tab: 1251 lab kit} shows the lab
kit contents for General Physics Laboratory I, and Table \ref{tab: 1261 lab kit} shows the lab
kit contents for General Physics Laboratory II.

\begin{table}
  \caption{\label{tab: 1261 lab kit} General Physics Laboratory II lab kit contents.}
  \begin{tabular}{clc}
    \hline\hline
    Quantity & Item & Investigation(s)\\
    \hline
    1 & \SI{100}{\ohm} resistor & 1 \\
    1 & \SI{330}{\ohm} resistor & 1 \\
    1 &  \SI{100}{\ohm} potentiometer & 1 \\
    1 & E10 light bulb holder & 1 \\
    1 & \SI{5}{V} E10 incandescent light bulb & 1 \\
    1 & Breadboard & 1, 2, 3 \\
    1 & Breadboard power supply & 1, 2, 3 \\
    1 & USB power supply cable & 1, 2, 3 \\
    1 & Jumper wire set & 1, 2 \\
    2 & Multimeters & 1, 2 \\
    1 & Mini screwdriver for multimeters & 1,2 \\
    5 & \SI{500}{mA} fusses for multimeters & 1,2 \\
    4 & Alligator clip leads & 1, 2 \\
    1 & \SI{1}{\mega\ohm} resistor & 2 \\
    2 & \SI{100}{\micro\farad} capacitor & 2 \\
    1 & \SI{5}{V}, \SI{650}{nm} laser module & 3 \\
    1 & Tape measure with cm scale & 3 \\            
    \hline\hline
  \end{tabular}
\end{table}

\section{Conclusion}
Discussion Points:
\begin{itemize}
  \item What have been our successes?  (Related to both Transformations)
  \item What have been our challenges?
  \item What plans do we have for online labs in the future?
\end{itemize}

\begin{acknowledgments}
  Redacted for review.% This work has been supported by the NSF (Award \# 1725655).
\end{acknowledgments}

\bibliography{onlinelabs.bib}

\end{document}